%\vfill\eject
\chapter{Basic concepts}
\label{basicchapter}

\section{Variables and regions}
\label{specialformsection}
\label{variablesection}

Any identifier that is not a syntactic keyword\index{keyword} (see
section~\ref{keywordsection}) may be used as a variable.\index{syntactic
keyword}\index{identifier}\mainindex{variable}  A variable may name a
location where a value can be stored.  A variable that does so is said
to be {\em bound} to the location.  The set of all such
bindings\mainindex{binding} in effect at some point in a program is
known as the {\em environment} in effect at that point.  The value
stored in the location to which a variable is bound is called the
variable's value.  By abuse of terminology, the variable is sometimes
said to name the value or to be bound to the value.  This is not quite
accurate, but confusion rarely results from this practice.

\todo{Define ``assigned'' and ``unassigned'' perhaps?}

\todo{In programs without side effects, one can safely pretend that the
variables are bound directly to the arguments.  Or:
In programs without \ide{set!}, one can safely pretend that the
variable is bound directly to the value. }

\vest Certain expression types are used to create new locations and to
bind variables to those locations.  The most fundamental of these {\em
binding constructs}\mainindex{binding construct} is the
\lambdaexp{}\index{\lambdaexp{}}, because all other binding constructs
can be explained in terms of \lambdaexp{}s.  The other binding
constructs are \ide{let}, \ide{let*}, \ide{letrec}, and \ide{do}
expressions (see sections~\ref{lambda}, \ref{letrec}, and \ref{do}).

%Note: internal definitions not mentioned here.

\vest Like Algol and Pascal, and unlike most other dialects of Lisp
except for Common Lisp, Scheme is a statically scoped language with
block structure.  To each place where a variable is bound in a program
there corresponds a \defining{region} of the program text within which
the binding is effective.  The region is determined by the particular
binding construct that establishes the binding; if the binding is
established by a \lambdaexp{}, for example, then its region
is the entire \lambdaexp{}.  Every reference to or assignment of a
variable refers to the binding of the variable that established the
innermost of the regions containing the use.  If there is no binding of
the variable whose region contains the use, then the use refers to the
binding for the variable in the top level environment, if any
(section~\ref{initialenv}); if there is no binding for the identifier,
it is said to be \defining{unbound}.\mainindex{bound}\index{top level
environment}

\todo{Mention that some implementations have multiple top level environments?}

\todo{Pitman sez: needs elaboration in case of {\tt(let ...)}}

\todo{Pitman asks: say something about vars created after scheme starts?
{\tt (define x 3) (define (f) x) (define (g) y) (define y 4)}
Clinger replies: The language was explicitly
designed to permit a view in which no variables are created after
Scheme starts.  In files, you can scan out the definitions beforehand.
I think we're agreed on the principle that interactive use should
approximate that behavior as closely as possible, though we don't yet
agree on which programming environment provides the best approximation.}


\section{True and false}

Any Scheme value can be used as a boolean value for the purpose of a
conditional test.  As explained in section~\ref{booleansection}, all
values count as true in such a test except for \schfalse{} and the empty
list, which count as false.  This report uses the word ``true'' to refer
to any Scheme value that counts as true, and the word ``false'' to refer
to any Scheme value that counts as false.
\mainindex{true} \mainindex{false}

\todo{Bartley: tighten this up.}


\section{External representations}
\label{externalreps}

An important concept in Scheme (and Lisp) is that of the {\em external
representation} of an object as a sequence of characters.  For example,
an external representation of the integer 28 is the sequence of
characters ``{\tt 28}'', and an external representation of a list consisting
of the integers 8 and 13 is the sequence of characters ``{\tt(8 13)}''.

The external representation of an object is not necessarily unique.  The
integer 28 also has representations ``{\tt 28.000}'' and ``{\tt\#x1c}'', and the
list in the previous paragraph also has the representations ``{\tt( 08 13
)}'' and ``{\tt(8 .\ (13 .\ ()))}'' (see section~\ref{listsection}).

Many objects have standard external representations, but some, such as
procedures, do not have standard representations (although particular
implementations may define representations for them).

An external representation may be written in a program to obtain the
corresponding object (see \ide{quote}, section~\ref{quote}).

External representations can also be used for input and output.  The
procedure \ide{read} (section~\ref{read}) parses external
representations, and the procedure \ide{write} (section~\ref{write})
generates them.  Together, they provide an elegant and powerful
input/output facility.

Note that the sequence of characters ``{\tt(+ 2 6)}'' is {\em not} an
external representation of the integer 8, even though it {\em is} an
expression evaluating to the integer 8; rather, it is an external
representation of a three-element list, the elements of which are the symbol
{\tt +} and the integers 2 and 6.  Scheme's syntax has the property that
any sequence of characters which is an expression is also the external
representation of some object.  This can lead to confusion, since it may
not be obvious out of context whether a given sequence of characters is
intended to denote data or program, but it is also a source of power,
since it facilitates writing programs such as interpreters and
compilers which treat programs as data (or vice versa).

The syntax of external representations of various kinds of objects
accompanies the description of the primitives for manipulating the
objects in the appropriate sections of chapter~\ref{initialenv}.
