% 1. Structure of the language

\chapter{Overview of Scheme}

\section{Semantics}
\label{semanticsection}

This section gives an overview of Scheme's semantics.  A
detailed informal semantics is the subject of
chapters~\ref{basicchapter} through~\ref{builtinchapter}.  For reference
purposes, section~\ref{formalsemanticssection} provides a formal
semantics of Scheme.

\vest Following Algol, Scheme is a statically scoped programming
language.  Each use of a variable is associated with a lexically
apparent binding of that variable.

\vest Scheme has latent as opposed to manifest types.  Types
are associated with values (also called objects) rather than with
variables.  (Some authors refer to languages with latent types as
weakly typed or dynamically typed languages.)  Other languages with
latent types are APL, Snobol, and other dialects of Lisp.  Languages
with manifest types (sometimes referred to as strongly typed or
statically typed languages) include Algol 60, Pascal, and~C.

\vest All objects created in the course of a Scheme computation, including
procedures and continuations, have unlimited extent.
No Scheme object is ever destroyed.  The reason that
implementations of Scheme do not (usually!)\ run out of storage is that
they are permitted to reclaim the storage occupied by an object if
they can prove that the object cannot possibly matter to any future
computation.  Other languages in which most objects have unlimited
extent include APL and other Lisp dialects.

\vest Implementations of Scheme are required to be properly tail-recursive.
This allows the execution of an iterative computation in constant space,
even if the iterative computation is described by a syntactically
recursive procedure.  Thus with a tail-recursive implementation,
iteration can be expressed using the ordinary procedure-call
mechanics, so that special iteration constructs are useful only as
syntactic sugar.

\vest Scheme procedures are objects in their own right.  Procedures can be
created dynamically, stored in data structures, returned as results of
procedures, and so on.  Other languages with these properties include
Common Lisp and ML. \todo{Rozas: Scheme had them first.}

\vest One distinguishing feature of Scheme is that continuations, which
in most other languages only operate behind the scenes, also have
``first-class'' status.  Continuations are useful for implementing a
wide variety of advanced control constructs, including non-local exits,
backtracking, and coroutines.  See section~\ref{continuations}.

\vest Arguments to Scheme procedures are always passed by value, which
means that the actual argument expressions are evaluated before the
procedure gains control, whether the procedure needs the result of the
evaluation or not.  ML, C, and APL are three other languages that always
pass arguments by value.
This is distinct from the lazy-evaluation semantics of SASL,
or the call-by-name semantics of Algol 60, where an argument
expression is not evaluated unless its value is needed by the
procedure.

\todo{Lisp's call by value should be explained more
accurately.  What's funny is that all values are references.}


\section{Syntax}

Scheme employs a parenthesized-list Polish notation to describe programs
and (other) data.  The syntax of Scheme, like that of most Lisp
dialects, provides for great expressive power, largely due to its
simplicity.  An important consequence of this simplicity is the
susceptibility of Scheme programs and data to uniform treatment by other
Scheme programs.  As with other Lisp dialects, the \ide{read} primitive
parses its input; that is, it performs syntactic as well as lexical
decomposition of what it reads.


\section{Notation and terminology}


\subsection{Essential and non-essential features}

\label{essentialsection}
It is required that every implementation of Scheme support
features that are marked as being \defining{essential}.  Features not
explicitly marked as essential are not essential.  Implementations are
free to omit non-essential features of Scheme or to add extensions,
provided the extensions are not in conflict with the language reported
here.


\subsection{Error situations and unspecified behavior}

\mainindex{error}
When speaking of an error situation, this report uses the phrase ``an
error is signalled'' to indicate that implementations must detect and
report the error.  If such wording does not appear in the discussion of
an error, then implementations are not required to detect or report the
error, though they are encouraged to do so.  An error situation that
implementations are not required to detect is usually referred to simply
as ``an error.''

\vest For example, it is an error for a procedure to be passed an argument that
the procedure is not explicitly specified to handle, even though such
domain errors are seldom mentioned in this report.  Implementations may
extend a procedure's domain of definition to include other arguments.

\vest If the value of an expression is said to be ``unspecified,'' then
the expression must evaluate to some object without signalling an error,
but the value depends on the implementation; this report explicitly does
not say what value should be returned.

\todo{Talk about unspecified behavior vs. unspecified values.}

\todo{Look at KMP's situations paper.}


\subsection{Entry format}

Chapters~\ref{expressionchapter} and~\ref{builtinchapter} are organized
into entries.  Each entry describes one language feature or a group of
related features, where a feature is either a syntactic construct or a
built-in procedure.  An entry begins with one or more header lines of the form

\noindent\pproto{\var{template}}{essential \var{category}}\unpenalty

if the feature is an essential feature, or simply

\noindent\pproto{\var{template}}{\var{category}}\unpenalty

if the feature is not an essential feature.

If \var{category} is ``\exprtype'', the entry describes an expression
type, and the header line gives the syntax of the expression type.
Components of expressions are designated by syntactic variables, which
are written using angle brackets, for example, \hyper{expression},
\hyper{variable}.  Syntactic variables should be understood to denote segments of
program text; for example, \hyper{expression} stands for any string of
characters which is a syntactically valid expression.  The notation
\begin{tabbing}
\qquad \hyperi{thing} $\ldots$
\end{tabbing}
indicates zero or more occurrences of a \hyper{thing}, and
\begin{tabbing}
\qquad \hyperi{thing} \hyperii{thing} $\ldots$
\end{tabbing}
indicates one or more occurrences of a \hyper{thing}.

If \var{category} is ``procedure'', then the entry describes a procedure, and
the header line gives a template for a call to the procedure.  Argument
names in the template are \var{italicized}.  Thus the header line

\noindent\pproto{(vector-ref \var{vector} \var{k})}{essential procedure}\unpenalty

indicates that the essential built-in procedure {\tt vector-ref} takes
two arguments, a vector \var{vector} and an exact non-negative integer
\var{k} (see below).  The header lines

\noindent%
\pproto{(append \vari{list} \varii{list})}{essential procedure}
\pproto{(append \var{list} \dotsfoo)}{procedure}\unpenalty

indicate that in all implementations, the {\tt append} procedure
must be defined to take two arguments, and some implementations will
extend it to take zero or more arguments.

\label{typeconventions}
It is an error for an operation to be presented with an argument that it
is not specified to handle.  For succinctness, we follow the convention
that if an argument name is also the name of a type, then this implies a
restriction on the type of that argument to the procedure.  For example,
the header line for {\tt vector-ref} given above dictates that first
argument to {\tt vector-ref} must be a vector.  The following naming
conventions also imply type restrictions:
\newcommand{\foo}[1]{\vr{#1}, \vri{#1}, $\ldots$ \vrj{#1}, $\ldots$}
$$
\begin{tabular}{ll}
\var{obj}&any object\\
\foo{z}&complex, real, rational, integer\\
\foo{x}&real, rational, integer\\
\foo{y}&real, rational, integer\\
\foo{q}&rational, integer\\
\foo{n}&integer\\
\foo{k}&exact non-negative integer\\
\end{tabular}
$$

\todo{Provide an example entry??}


\subsection{Evaluation examples}

The symbol ``\evalsto'' used in program examples should be read
``evaluates to.''  For example,

\begin{scheme}
(* 5 8)      \ev  40%
\end{scheme}

means that the expression {\tt(* 5 8)} evaluates to the object {\tt 40}.
Or, more precisely:  the expression given by the sequence of characters
``{\tt(* 5 8)}'' evaluates, in the initial environment, to an object
that may be represented externally by the sequence of characters ``{\tt
40}''.  See section~\ref{externalreps} for a discussion of external
representations of objects.


\todo{\subsection{Naming conventions}}
	
\todo{Terms that need defining: thunk, command, mutation procedure,
(what else?).  The {\tt?}, {\tt!}, and {\tt->} conventions should be
explained somewhere.}

\todo{Just say somewhere that the value of procedures whose names end in
{\tt!}\ is unspecified.}
